### 2024-12-19 NLP/exercises.tex
Minor updates.

\subject{Natural Language Processsing}
\semester{Winter 2024}
\author{Leopold Lemmermann}

@@

\setcounter{section}{2022}\sheet{1st exam}

@@

\setcounter{section}{2021}\sheet{1st exam}

@@

\begin{exercise}{Information retrieval metrics}\label{ex:ir-metrics}
  Two search engines have produced the following results ($+$ = relevant, $-$ = irrelevant).
  \begin{itemize}
    \item \textbf{Engine A}: $+, -, +, -, +, +, -, +, +, +$
    \item \textbf{Engine B}: $+, +, -, +, +, -, +, +, -, +$
  \end{itemize}

  \begin{enumerate}
    \item Calculate P@5, P@10 and MAP for both engines.
    \item How do these metrics help evaluate the performance of IR systems?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{Engine A}:
          \begin{itemize}
            \item $P@5 = \frac{3}{5} = 0.6$
            \item $P@10 = \frac{7}{10} = 0.7$
            \item $MAP = \frac{1}{3} \left( \frac{1}{1} + \frac{2}{4} + \frac{3}{5} \right) = 0.5666$
          \end{itemize}
          \item \textbf{Engine B}:
          \begin{itemize}
            \item $P@5 = \frac{4}{5} = 0.8$
            \item $P@10 = \frac{7}{10} = 0.7$
            \item $MAP = \frac{1}{3} \left( \frac{1}{1} + \frac{2}{4} + \frac{3}{7} \right) = 0.5476$
          \end{itemize}
        \end{itemize}
      \item These metrics help evaluate IR systems by measuring precision at different cutoff points and the average precision across all relevant documents.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{N-gram analysis}: Compare the frequency of character n-grams against language-specific profiles.
          \item \textbf{Machine learning models}: Train classifiers (e.g., SVM, neural networks) using labeled text data for various languages.
          \item \textbf{Lexicon-based methods}: Use dictionaries to match input text to language-specific words.
          \item \textbf{Hybrid approaches}: Combine n-gram, lexicon, and machine learning techniques for better accuracy.
        \end{itemize}
      \item \begin{itemize}
          \item \textbf{String matching}: Use algorithms to find exact matches between documents.
          \item \textbf{Semantic similarity}: Apply embeddings to detect paraphrased or rephrased content.
          \item \textbf{Citation analysis}: Check for missing or incorrect citations.
          \item \textbf{Fingerprinting}: Generate hashes of text fragments and compare across sources.
        \end{itemize}
      \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Expanding queries}: Use synonyms, related terms, or stemming to match more results.
        \item \textbf{Reducing filters}: Remove restrictive criteria (e.g., date, location).
        \item \textbf{Boosting document recall}: Index more documents or reduce relevance thresholds.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{enumerate}
        \item \begin{itemize}
            \item \textbf{Customer feedback}: Analyze reviews for product improvements.
            \item \textbf{Social media monitoring}: Detect trends and public sentiment.
            \item \textbf{Political analysis}: Gauge public opinion on policies or events.
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Factual sentiment analysis}: Focuses on extracting objective sentiment (e.g., product ratings).
            \item \textbf{Opinionated analysis}: Identifies subjective opinions (e.g., customer reviews).
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Supervised learning}: Uses labeled data to train machine learning models.
            \item \textbf{Lexicon-based approaches}: Use sentiment dictionaries to determine polarity.
          \end{itemize}
        \item Extend lexicons by identifying synonyms, antonyms, or similar words in distributed word embeddings like Word2Vec or GloVe.
      \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Factoid questions}: Require specific factual answers (e.g., "Who is the president of the U.S.?").
        \item \textbf{List questions}: Require a list of items (e.g., "List countries in the EU").
        \item \textbf{Definition questions}: Seek explanations or definitions (e.g., "What is AI?").
    \end{itemize}
  \end{solution}
\end{exercise}

@@

\setcounter{section}{2020}\sheet{1st exam}

>>>>

\subject{Natural Language Processing}
\semester{Winter 2024}
\author{Leopold Lemmermann}

@@

\setcounter{section}{2022}\sheet{Exam}

@@

\setcounter{section}{2021}\sheet{Exam}

@@

\begin{exercise}{Information retrieval metrics}\label{ex:ir-metrics}
  Two search engines have produced the following results ($+$ = relevant, $-$ = irrelevant).
  \begin{itemize}
    \item \textbf{Engine A}: $+-+-+\quad+-+++$
    \item \textbf{Engine B}: $---+-\quad-++-+$
  \end{itemize}

  \begin{enumerate}
    \item Calculate P@5, P@10 and MAP@5 for both engines.
    \item How do these metrics help evaluate the performance of IR systems?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{Engine A}:
            \begin{itemize}
              \item $P@5 = \frac{3}{5} = .6$
              \item $P@10 = \frac{7}{10} = .7$
              \item $MA@5 = (\frac{1}{1} + \frac{2}{3} + \frac{3}{5}) / 3 = .756$
            \end{itemize}
          \item \textbf{Engine B}:
            \begin{itemize}
              \item $P@5 = \frac{1}{5} = .2$
              \item $P@10 = \frac{4}{10} = .4$
              \item $MAP@5 = (\frac{1}{4}) / 1 = .25$
            \end{itemize}
        \end{itemize}
      \item These metrics help evaluate IR systems by measuring precision at different cutoff points and the average precision across all relevant documents.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{N-gram analysis}: Compare the frequency of character n-grams against language-specific profiles.
          \item \textbf{Machine learning models}: Train classifiers (e.g., SVM, neural networks) using labeled text data for various languages.
          \item \textbf{Lexicon-based methods}: Use dictionaries to match input text to language-specific words.
        \end{itemize}
      \item \begin{itemize}
          \item \textbf{String matching}: Use algorithms to find exact matches between documents.
          \item \textbf{Semantic similarity}: Apply embeddings to detect paraphrased or rephrased content.
          \item \textbf{Fingerprinting}: Generate hashes of text fragments and compare across sources.
        \end{itemize}
      \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Reducing filters}: Remove restrictive criteria (e.g., date, location).
        \item \textbf{Boosting document recall}: Index more documents or reduce relevance thresholds.

    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{enumerate}
        \item \begin{itemize}
            \item \textbf{Customer feedback}: Analyze reviews for product improvements.
            \item \textbf{Political analysis}: Gauge public opinion on policies or events.
            \item \textbf{Stock market prediction}: Predict stock prices based on sentiment in financial news.
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Factual sentiment analysis}: Focuses on extracting objective sentiment (e.g., product ratings).
            \item \textbf{Opinionated analysis}: Identifies subjective opinions (e.g., customer reviews).
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Supervised learning}: Uses labeled data to train machine learning models.
            \item \textbf{Lexicon-based approaches}: Use sentiment dictionaries to determine polarity.
          \end{itemize}
        \item Extend lexicons by identifying synonyms, antonyms, or similar words in distributed word embeddings like Word2Vec or GloVe.
      \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Factoid questions}: Require specific factual answers (e.g., "Who is the president of the U.S.?").
        \item \textbf{List questions}: Require a list of items (e.g., "What countries are in the EU?").
        \item \textbf{Definition questions}: Seek explanations or definitions (e.g., "What is AI?").
    \end{itemize}
  \end{solution}
\end{exercise}

@@

\setcounter{section}{2020}\sheet{Exam}
